Spin canonical divisors of log stacky curves

### 2015-06-29 stacky_write_up/spin_cring.tex
write monomial ordering ssec of bg

\begin{defn}
We equip the polynomial ring $k[x_1, \ldots, x_n]$ with the
(weighted graded) reverse lexicographic order \textbf{grevlex}
$\prec$: if 
\[
	x^{\vec{m}} = \prod_{d,i} x_{d,i}^{m_{d,i}}
\]

\noindent
and $x^{\vec{n}}$ are monomials in $k[x]_{\vec{a}}$, then $x^{\vec{m
}} \succ x^{\vec{n}}$ if and only if either
\begin{equation}
\vec{a}\cdot\vec{m} = \sum_{d,i} dm_{d,i}  > \vec{a} \cdot \vec{n}
\end{equation}
or
\begin{equation} \label{eqn:grevlexdef}
\text{$\vec{a}\cdot\vec{m} = \vec{a} \cdot \vec{n}$ and the last (right-most) nonzero entry in $\vec{m}-\vec{n}$ is negative.}
\end{equation}
It is important to note that in \eqref{eqn:grevlexdef}, the ordering of the variables matters: it corresponds to a choice of writing the exponents of a monomial as a vector.  A common choice for us will be
\begin{equation} \label{eqn:xdOrder}
 x_{1,1}^{m_{1,1}} \cdots x_{D,a_D}^{m_{D,a_D}} \leftrightarrow (m_{D,1},\dots,m_{D,a_D},\ldots,m_{1,1},\dots,m_{1,a_1}) 
 \end{equation}
in which case we have $x_{2,1} \succ x_{1,2}^2$
and $x_{5,1} \succ x_{1,6}^2 x_{3,2}$, etc.
We indicate this ordering in the presentation of the ring, e.g.~for the above we would write
\[ k[x]_{\vec{a}}= k[x_{1,1},\dots,x_{1,a_1},\dots,x_{D,1},\dots,x_{D,a_D}]. \]
In this way, our relations write generators in larger degree in terms of those in smaller degree (``later generators in terms of earlier generators''), which gives the most natural-looking canonical rings to our eyes.
\end{defn}

For a homogeneous polynomial $f \in k[x]_{\vec{a}}$, we define the \textbf{initial term} $\initial_{\prec}(f)$ to be the monomial in the support of $f$ that is largest with respect to $\prec$, and we define $\initial_{\prec}(0)=0$.  Let $I \subseteq k[x]_{\vec{a}}$ be a homogeneous ideal.  We define the \textbf{initial ideal} $\initial_{\prec}(I)$ to be the ideal generated by $\{\initial_{\prec}(f) : f \in I\}$.  A \textbf{Gr\"obner basis} (also called a \textbf{standard basis}) for $I$ is a set $\{f_1,\dots,f_n\} \subset I$ such that 
\[ \langle \initial_{\prec}(f_1),\dots,\initial_{\prec}(f_n) \rangle = \initial_{\prec}(I). \]

@@

We also introduce a term ordering that is well-suited for inductive arguments: the \emph{block} term order.  
In our inductive arguments, we will often have the following setup: an inclusion $R \supset R'$ of canonical rings such that $R'$ is generated by elements $x_{i,d}$ and $R$ is generated over $R$ by elements $y_j$.  It is natural, therefore, to consider term orders which treat these sets of variables separately.  More formally, we make the following definition.

Let $k[x]_{\vec{a}}$ and $k[y]_{\vec{b}}$ be weighted polynomial rings with term orders $\prec_x$ and $\prec_y$, respectively, and consider $k[y,x]_{\vec{b},\vec{a}} = k[y]_{\vec{b}} \otimes_k k[x]_{\vec{a}}$ the common weighted polynomial ring in these two sets of variables $y,x$.

\begin{defn}
The \textbf{(graded) block} (or \textbf{elimination}) term ordering on $k[y,x]_{\vec{b},\vec{a}}$ is defined as follows: we declare
  \[ y^{\vec{m}}x^{\vec{n}}  \succ y^{\vec{m}'}x^{\vec{n}'}\]  
  if and only if
  \begin{enumerate}
  \item[(i)] $\deg y^{\vec{m}} x^{\vec{n}}  > \deg y^{\vec{m}'} x^{\vec{n}'} $, or
  \item[(ii)] $\deg y^{\vec{m}}x^{\vec{n}} = \deg y^{\vec{m}'} x^{\vec{n}'} $ and 
    \begin{enumerate}
    \item $y^{\vec{m}}\succ_y y^{\vec{m}'}$ or
    \item $y^{\vec{m}} =  y^{\vec{m}'}$ and $x^{\vec{n}} \succ_x x^{\vec{n}'}$.
    \end{enumerate}
  \end{enumerate}
\end{defn}

The block ordering is indeed a term order: the displayed inequalities directly give that any two monomials are comparable, and the inequalities are visibly stable under multiplication by a monomial.  One can similarly define an \textbf{iterated (graded) block} term ordering for any finite number of weighted polynomial rings.

The block ordering is the most suitable ordering for the structure of $R$ as an $R'$-algebra

background on Gr\"{o}bner bases
Notation for Gr\"{o}bner bases and orderings (include grlex, grevlex, and block order, as defined in \cite[Definition 8.1.1]{vzb:stacky})

@@

\begin{defn}
\label{defn:lower-ideal}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $I_{L,d}$ be the $d$th graded piece of $I_L$ and define
$$J_{L,d} = \sum_{j = 1}^{d- 1}k[x_1, \ldots, x_n]_j \cdot I_{L,d-j}.$$
\end{defn}

@@

\noindent
so $R'$ must have some generators $x_{3, 1}, x_{5, 1}, x_{7, 1},
x_{7, 2}$ with $x_{i,j} = H^0(\sx',iK_{\sx'})$.

>>>>

First, we give a definition of the graded degree of a monomial
in a graded polynomial ring.

\begin{defn}
\label{defn:monomial-degree}
Let $k[x_{d_1, 1}$, $\ldots$, $x_{d_n, n}]$ be a graded polynomial
ring and let $\alpha := \prod_{i = 1}^{n} x_{d_i, i}^{f_i} \in
k[x_{d_1, 1}$, $\ldots$, $x_{d_n, n}]$ be a monomial. Then
we define

\[
	\deg \alpha := \sum_{i = 1}^{n} d f_i
\]
\end{defn}

Now define the (graded) lexicographic order \textbf{grlex}
$\prec_{\text{grlex}}$ on a polynomial ring $k[x_{d_1, 1}, \ldots,
x_{d_n, n}]$ to be the monomial ordering such that:

\begin{defn}
\label{defn:grlex}
If $\alpha := \prod_{i = 1}^{n} x_{d_i, i}
^{f_i}$ and $\beta := \prod_{i = 1}^{n} x_{d_i, i}^{f_i'}$ are
monomials in $k[x_{d_1, 1}$, $\ldots$, $x_{d_n, n}]$, then $\alpha
\succ_{\text{grlex}} \beta$ if either $\alpha$ has larger degree:
\begin{equation}
	\deg \alpha := \sum_{i = 1}^{n} d f_i  > \sum_{i = 1}^{n} d f_i' =: \deg \beta
\end{equation}
or $\alpha$ is larger using the lexicographical ordering event of equal
degrees:

\begin{equation}
\label{eqn:grlex-tie}
	\deg \alpha = \deg \beta \text{ and }
	\alpha \succ_{\text{lex}} \beta
\end{equation}
\end{defn}

We can equip the polynomial ring $k[x_{d_1, 1}, \ldots, x_{d_n, n}]$
with the (weighted graded) reverse lexicographic order
\textbf{grevlex} $\prec_{\text{grevlex}}$ as follows:

\begin{defn}
\label{defn:grevlex}
If $\alpha := \prod_{i = 1}^{n} x_{d_i, i}
^{f_i}$ and $\beta := \prod_{i = 1}^{n} x_{d_i, i}^{f_i'}$ are
monomials in $k[x_{d_1, 1}$, $\ldots$, $x_{d_n, n}]$, then $\alpha
\succ_{\text{grevlex}} \beta$ if either $\alpha$ has larger degree:

\begin{equation}
	\deg \alpha := \sum_{i = 1}^{n} d f_i  > \sum_{i = 1}^{n} d f_i' =: \deg \beta
\end{equation}
or $\alpha$ has a smaller right-most entry in the event of equal
degrees:

\begin{equation}
\label{eqn:grevlex-tie}
	\deg \alpha = \deg \beta \text{ and }	f_i - f_i' < 0 \text{ for
the	largest } i \text{ such that } f_i \neq f_i'
\end{equation}
\end{defn}

\begin{rem}
Note that the ordering of the variables matters in both
Equation ~\ref{eqn:grlex-tie} and Equation ~\ref{eqn:grevlex-tie}.
\end{rem}

Our inductive arguments in Section ~\ref{sec:induction} will
usually have an inclusion $R \supset R'$ of log spin canonical
rings such that $R'$ is generated by elements $x_{d_i, i}$ and $R$
is generated over $R'$ by elements $y_j$. In these cases, it is
natural to consider term orders which treat these sets of variables
separately. We now present the \textbf{block term order} introduced
by Voight and Zureick-Brown \cite[Section 8.1]{vzb:stacky}.

\begin{defn}
\label{defn:block-order}
Let $k[x_{d_1, 1}, \ldots, x_{d_n, n}]$ and $k[y_{c_1, 1}, \ldots,
y_{c_m, m}]$ be weighted polynomial rings with existing term
orders $\prec_x$ and $\prec_y$, respectively. Let $\alpha :=
\prod_{i = 1}^{n} x_{d_i, i}^{f_i}$ $\prod_{j = 1}^{m} y_{c_j, j}
^{h_i}$ and $\beta := \prod_{i = 1}^{n} x_{d_i, i}^{f_i'}
\prod_{j = 1}^{m} y_{c_j, j}^{h_i'}$ be monomials in $k[x_{d_1, 1},
\ldots, x_{d_n, n}] \otimes$ $k[y_{d_1', 1}, \ldots, y_{d_m', m}]$.
Let $\alpha_x := \prod_{i = 1}^{n} x_{d_i, i}^{f_i}$ be the
part of $\alpha$ in $k[x_{d_1, 1}, \ldots, x_{d_n, n}]$ and
likewise with $\alpha_y$, $\beta_x$, and $\beta_y$.

In the \textbf{(graded) block} (or \textbf{elimination}) term
ordering on $k[x_{d_1, 1}, \ldots, x_{d_n, n}] \otimes k[y_{d_1', 1},
\ldots, y_{d_m', m}]$, we define $\alpha \succ \beta$ if
  \begin{enumerate}
  \item[(i)] $\deg \alpha  > \deg \beta$ or
  \item[(ii)] $\deg \alpha = \deg \beta$ and 
    \begin{enumerate}
    \item $\alpha_y \succ_y \beta_y$ or
    \item $\alpha_y =  \beta_y$ and $\alpha_x \succ_x \beta_x$
    \end{enumerate}
  \end{enumerate}
\end{defn}

\todo{write below part}
For a homogeneous polynomial $f \in k[x]_{\vec{a}}$, we define the \textbf{initial term} $\initial_{\prec}(f)$ to be the monomial in the support of $f$ that is largest with respect to $\prec$, and we define $\initial_{\prec}(0)=0$.  Let $I \subseteq k[x]_{\vec{a}}$ be a homogeneous ideal.  We define the \textbf{initial ideal} $\initial_{\prec}(I)$ to be the ideal generated by $\{\initial_{\prec}(f) : f \in I\}$.  A \textbf{Gr\"obner basis} (also called a \textbf{standard basis}) for $I$ is a set $\{f_1,\dots,f_n\} \subset I$ such that 
\[ \langle \initial_{\prec}(f_1),\dots,\initial_{\prec}(f_n) \rangle = \initial_{\prec}(I). \]

@@

@@

\begin{defn}
\label{defn:lower-ideal}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $I_{L, d}$ be the $d$th graded piece of $I_L$ and define
$$J_{L,d} = \sum_{j = 1}^{d- 1}k[x_1, \ldots, x_n]_j \cdot I_{L,d-j}.$$
\end{defn}

@@

\noindent
so $R'$ must have some generators $x_{3, 1}, x_{5, 1}, x_{7, 1},
x_{7, 2}$ with $x_{i,j} \in H^0(\sx', i K_{\sx'})$.
